### 2015-01-03 Studies/2013_PG_County/Report/PG_Co_Fire.tex
PG Co: minor update to report

November 2014}}

@@

\chapter*{\centering Abstract}
Fire Dynamics Simulator (FDS), which is a fire model that is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on February 24, 2012, within a single-story plus basement, single-family residential structure in Riverdale Heights, MD, that resulted in the injury of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the front door was opened, the interior stairwell, aided by strong winds blowing into the basement, acted as a chimney for hot gases in the basement to flow towards regions of lower pressure, out through the front door. The temperature of the gases in the interior stairwell was estimated to be in excess of 400~$^{\circ}$C (750~$^{\circ}$F). Two firefighters were located in the flow path between the interior stairwell door and the front door of the structure, were exposed to these elevated temperatures, and suffered serious burn injures. Four additional firefighters were injured during fire operations on the first floor.
\pagenumbering{gobble}

@@

On February 24, 2012, a fire in residential structure injured seven Prince George's County firefighters. NIST examined the fire dynamics of this incident at the request of the Prince George's Country Fire/Emergency Medical Services (EMS) Department. Computer simulations of the fire incident were conducted using the NIST Fire Dynamics Simulator~\cite{FDS_Users_Guide} and Smokeview~\cite{Smokeview_Users_Guide}  software to provide insight into the fire development and thermal conditions that likely existed in the residence during the fire. The specific objectives of the simulations detailed in this report are: 
\begin{enumerate}
\item To examine the effect of the flow path (including temperature, pressure, and fire conditions) in this multi-story residential structure using physics-based calculations.
\item To provide visualizations of the fire behavior that are representative of the conditions that members of the Prince George's Fire Department likely experienced during the course of their interior operations.
\end{enumerate}

@@

The analysis of the simulation results is focused on the impact of ventilation. Specifically on the  interior conditions of the first floor following the establishment of a flow path through door control and how wind impacted these conditions. Based on the goals of the analysis and the information collected, several engineering approximations were made. For example, heavy smoke conditions were reported at the front of the house at the time of arrival~\cite{PGCounty2013}; however, the exact ignition time, fire growth rate, and heat release rate (HRR) of the fire were not known. The ignition time and growth rate were estimated to reduce the amount of required calculation time while adhering to event timeline (Table~\ref{tab:fire_info}) and documented observations. The HRR of the fire was estimated using documentation of the potential fuel load within the structure.

\section{Geometry}
\label{geom}
The structure involved in this fire incident was a one story, single-family structure with a basement. The structure was platform-framed wood construction with a gable roof. \footnote{Platform-framed construction means that floor joists are completely covered with a sub-floor. This forms a platform for which walls are built from.} The structure had a first floor footprint of 9.14~m (30~ft) by 7.77~m (25.5~ft) with the basement being slightly smaller, 9.14~m (30~ft) by 7.16~m (23.5~ft), as show in Fig.~\ref{fig:simp_geom}. This resulted in a first floor area of 71~m$^2$ (764~ft$^2$) and a basement floor area of 65~m$^2$ (700~ft$^2$). The interior walls were composed of gypsum board over insulated wood framing.~\cite{PGCounty2013}. There was a single flight of stairs in the center of the structure which connected the first floor to the basement.

The structure was built on a sloped landscape such that on the front side of the structure the first floor was located above street level but the basement was below street level (Fig.~\ref{fig:slope}). On the rear side of the structure, the slope allowed the basement to be a walkout. Therefore, access to the structure could be made on the front side of the first floor and the rear side of the basement.

@@

Two firefighters were located in the flow path between the basement and the doors on the front side of the structure. One firefighter was able to break a window to escape the hazardous conditions and after a call for assistance the second firefighter was removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where they were treated for their injuries. Four other firefighters were injured on-scene and transported to the hospital for treatment.

>>>>

January 2015}}

@@

\chapter*{\centering Abstract}
Fire Dynamics Simulator (FDS), which is a fire model that is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on February 24, 2012, within a single-story, single-family residential structure in Riverdale Heights, MD, that resulted in the serious injury of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the front door was opened the interior stairwell, aided by strong winds blowing into the basement, acted as a chimney for hot gases in the basement to flow towards regions of lower pressure, out through the front door. The temperature of the gases in the interior stairwell was estimated to be in excess of 400~$^{\circ}$C (750~$^{\circ}$F). Two firefighters were located in the flow path between the interior stairwell door and the front door of the structure, were exposed to these elevated temperatures, and suffered serious burn injures. Five additional firefighters were injured during fire operations on the first floor.
\pagenumbering{gobble}

@@

On February 24, 2012, a fire in residential structure injured seven Prince George's County firefighters. NIST examined the fire dynamics of this incident at the request of the Prince George's Country Fire/Emergency Medical Services (EMS) Department. Computer simulations of the fire incident were conducted using the NIST Fire Dynamics Simulator~\cite{FDS_Users_Guide} and Smokeview~\cite{Smokeview_Users_Guide}  software to provide insight into the fire development and thermal conditions that likely existed in the residence during the fire. The specific objectives of the simulations detailed in this report are: 
\begin{enumerate}
\item To examine the effect of the flow path (including temperature, pressure, and fire conditions) in this single-story residential structure using physics-based calculations.
\item To provide visualizations of the fire behavior that are representative of the conditions that members of the Prince George's Fire Department likely experienced during the course of their interior operations.
\end{enumerate}

@@

The analysis of the simulation results is focused on the impact of ventilation. Specifically on the interior conditions of the first floor following the establishment of a flow path through door control and how wind impacted these conditions. Based on the goals of the analysis and the information collected, several engineering approximations were made. For example, heavy smoke conditions were reported at the front of the house at the time of arrival~\cite{PGCounty2013}; however, the exact ignition time, fire growth rate, and heat release rate (HRR) of the fire were not known. The ignition time and growth rate were estimated to reduce the amount of required calculation time while adhering to the event timeline (Table~\ref{tab:fire_info}) and documented observations. The HRR of the fire was estimated using documentation of the potential fuel load within the structure and relevant literature.

\section{Geometry}
\label{geom}
The structure involved in this fire incident was a one story, single-family structure with a basement. The structure was platform-framed wood construction with a gable roof. \footnote{Platform-framed construction means that floor joists are completely covered with a sub-floor. This forms a platform for which walls are built from.} The structure had a first floor footprint of 9.14~m (30~ft) by 7.77~m (25.5~ft) with the basement being slightly smaller, 9.14~m (30~ft) by 7.16~m (23.5~ft), as show in Fig.~\ref{fig:simp_geom}. This resulted in a first floor area of 71~m$^2$ (764~ft$^2$) and a basement floor area of 65~m$^2$ (700~ft$^2$). The interior walls were composed of gypsum board over insulated wood framing~\cite{PGCounty2013}. There was a single flight of stairs in the center of the structure which connected the first floor to the basement.

The structure was built on a sloped landscape such that on the front side of the structure the first floor was located at street level but the basement was below street level (Fig.~\ref{fig:slope}). On the rear side of the structure, the slope allowed the basement to be a walkout. Therefore, access to the structure could be made on the front side to the first floor and on the rear side to the basement.

@@

Two firefighters were located in the flow path between the basement and the doors on the front side of the structure. One firefighter was able to break a window to escape the hazardous conditions and after a call for assistance the second firefighter was removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where they were treated for their injuries. Five other firefighters were injured on-scene and transported to the hospital for treatment.
